---
title: "MADS Project"
output: word_document
date: "2023-11-27"
---

Data and system set-up
```{r}
##workingdrivelaptop
##setwd('/Users/duynguyen/Library/CloudStorage/OneDrive-UniversityofSanDiego/ADS-500B/Project') 

##workingdrivedesktop
setwd('/Users/usd-msads/Documents/Documents - DUY’s Mac mini/GitHub/500B Project/') 

##map data
df <- read.table('house_sales.csv', header = TRUE, sep=",")

#create dataframe 2 to process and cleanup values
df2<- df

#ensure libraries are ran
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
library(reshape2)
library(corrplot)
```

Data Definitions
This dataset contains details of house sales in neighborhoods of King County, Washington. 
It includes data for houses sold between May 2014 and May 2015.  

Origin: Kaggle
File Type: CSV
File Size: 2.5 MB
Modeling Problem: Regression
Missing Values: Yes

Attributes:
id - a numerical identity field for each house sale
date - date when the house is sold
price - house sale price amount in US dollars (dependent variable)
bedrooms - number of bedrooms in the house
bathrooms - number of bathrooms in the house
sqft_living - living area in square feet
sqft_lot - total lot area in square feet
floors - number of floors in the house
waterfront - binary field, 1 if the house has a waterfront, 0 for no waterfront
view - binary field, 1 if the house has a view, 0 for no view
condition - a number denoting the condition of the house
grade - a number denoting the grade of the house
sqft_basement - area in square feet if the house has a basement
yr_built - in which year is the house built
yr_renovated - year when a house is renovated, 0 if not renovated
zipcode - house location zipcode in the county
lat - latitude cooridnate 
long - longitude coordinate


Run Exploratory Data Analysis to preprocess and clean up data
1. Run summary to look at columns that have NA values and evaluate descriptive statistics
````{r}
summary(df2)
##create data frame of NA value to revaluate details and determine relationships with other columns so we can determine appropirate cleanup method
na_df2 <- df2[rowSums(is.na(df2)) > 0,]
head(na_df2, 10)
```

1.1 The following columns have NAs; bedrooms, bathrooms, sqft_living, sqft_lot.  The following steps have been taken to clear the dataset of NAs

1.1.1 Bedrooms and bathrooms - replace 33 with 3 bedrooms and NA with mean bedrooms
```{r}
df2$bedrooms[df2$bedrooms == 33] <- 3
df2$bedrooms[is.na(df2$bedrooms)]<-mean(df2$bedrooms,na.rm=TRUE)
df2$bathrooms[is.na(df2$bathrooms)]<-mean(df2$bathrooms,na.rm=TRUE)
```

1.1.2 Remove NA values in sqft_living and sqft_lot, descritive statistics not normally distributed enough to use mean
```{r}
#store removal of NA data in sqft_living and sqft_lot
df2<- df2[!is.na(df2$sqft_living),]
df2<- df2[!is.na(df2$sqft_lot),]
```

1.2 Evaluate data set for extreme outliers and remove from data set
```{r}
#create dataframe to evalute house features
bp_df2 <- df2[, c('bedrooms','bathrooms','floors','grade')]
boxplot(bp_df2)
```

[include evaluation of box plots]

```{r}
#create dataframe to evalute house feature
bp2_df2 <- df2[, c('sqft_living','sqft_above','sqft_basement','sqft_living15')]
boxplot(bp2_df2)
```

[include evaluation of box plots]

```{r}
#create dataframe to evalute house feature
bp3_df2 <- df2[, c('sqft_lot', 'sqft_lot15')]
boxplot(bp3_df2)
```

[From sqft_lot remove values greater than 1500000]

```{r}
#Remove outler record from sqft_lot that is greater than 1500000 to create final data frame for analysis
df_f <- subset(df2,df2$sqft_lot < 1500000)
head(df_f, 10)
```
```{r}
#Descriptive Statistics
summary(df_f)
```

2. Descriptive Statistics of columns

Datatypes

date - discrete data
price - discrete data
bedrooms - discrete data
bathrooms - discrete data
sqft_living - continuous data
sqft_lot - continuous data
floors - discrete data
waterfront - discrete data
view - categorical data
condition - ordinal data
grade - ordinal data
sqft_basement - continuous data
yr_built - discrete data
yr_renovated - discrete data
zipcode - categorical
lat - Continuous data
long - Continuous data

```{r}
#create data frame of continuous and discrete data to run correlation test
df_f_corr <- df_f[, c('price', 'bedrooms', 'bathrooms', 'sqft_living', 'sqft_lot', 'floors', 'waterfront', 'sqft_basement', 'yr_built', 'yr_renovated', 'zipcode', 'lat', 'long')]
df_f_corr <- scale(df_f_corr)
cor(df_f_corr)
```
```{r}
#create a model with correlated values price and sqft_living
model<- lm(price ~ sqft_living, data=df_f)
print(model)
```



```{r}
#create scatter plot to look at linear relationship
ggplot(df_f,aes(x=sqft_living, y = price)) + geom_point() + stat_smooth()
```

```{r}
#create scatter plot to look at linear relationship, straight line
ggplot(df_f,aes(x=sqft_living, y = price)) + geom_point() + stat_smooth(method=lm)
```
```{r}
summary(model)
```
  [value used the model are statistically significant P >.000, adjust r-square value is .4926, ~50 of the dependent variable explains the independent variable ]
  
```{r}
#create a model with correlated values price and sqft_living
model2<- lm(price ~ sqft_living + bedrooms, data=df_f)
summary(model2)
```